---
title: "project1"
author: "Peiyu Chen"
date: "2024-09-07"
output: html_document
---


## check package and data


```{r  message=FALSE}

## Install the tidyverse package if you don't have it
if (!require("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
}

## Install the remotes package if you don't have it
if (!require("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
## Install the tidytuesdayR package (from GitHub) if you don't have it
if (!require("tidytuesdayR", quietly = TRUE)) {
    remotes::install_github("dslc-io/tidytuesdayR")
}

##to aviod redownloading data:

library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
  dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
  url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
  chocolate <- readr::read_csv(url_csv)
  
  saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

glimpse(chocolate)

```


## Part 1: Explore data


1. Make a histogram of the rating scores to visualize the overall distribution scores. 
Change the number of bins from default to 10, 15, 20, and 25. 
Pick on the one that you think looks the best.Explain what the difference is when you changes the number of bins and explain why you picked the one you did. 


```{r message=FALSE}
hist(chocolate$rating, main = "#bins=default")
hist(chocolate$rating, breaks=10, main = "#bins=10")
hist(chocolate$rating, breaks=15, main = "#bins=15")
hist(chocolate$rating, breaks=20, main = "#bins =20")
hist(chocolate$rating, breaks=25, main = "#bins =25")

```

2. Consider the countries where the beans originated from. How many reviews come from each country of bean origin?

```{r}
beans <- filter(chocolate, company_location == country_of_bean_origin)
count(group_by(beans,country_of_bean_origin))
```
 
 
 3. What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? 
 
 For this same set of reviews, also calculate 
 
 (1) the total number of reviews and 
 
 (2) the standard deviation of the rating scores. 
 
 Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.
 

```{r}
ecuador <- chocolate %>% group_by(country_of_bean_origin)%>% filter(country_of_bean_origin == "Ecuador")

summary <- ecuador %>%
  summarise (
    mean = mean(rating),
    sd = sd(rating),
    total = n()
  )
print(summary)
```


4. Which company location makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?


```{r}

maxecuador <- ecuador %>% 
  group_by(company_location)%>%
  summarise(mean= mean(rating)) %>%
  arrange(desc(mean))%>%
  slice(1)
print(maxecuador)

```

5. Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?


```{r}
high3 <- chocolate %>% group_by(country_of_bean_origin) %>%
  summarise(mean=mean(rating))%>%
  arrange(desc(mean))%>%
  slice(1:3)
print(high3)
```


6. Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?

```{r}
high3_10 <- chocolate %>% group_by(country_of_bean_origin) %>%
  filter(n() >=10)%>%
  summarise(mean=mean(rating))%>%
  arrange(desc(mean))%>%
  slice(1:3)

print(high3_10)
```

7. For this last part, let’s explore the relationship between percent chocolate and ratings.

Use the functions in dplyr, tidyr, and lubridate to perform the following steps to the chocolate dataset:

    a. Identify the countries of bean origin with at least 50 reviews. Remove reviews from countries are not in this list.

    b. Using the variable describing the chocolate percentage for each review, create a new column that groups chocolate percentages into one of four groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90% 
    (Hint: check out the substr() function in base R and the case_when() function from dplyr – see example below).
    
    c. Using the new column described in #2, re-order the factor levels (if needed) to be starting with the smallest percentage group and increasing to the largest percentage group (Hint check out the fct_relevel() function from forcats).
    
    d. For each country, make a set of four side-by-side boxplots plotting the groups on the x-axis and the ratings on the y-axis. These plots should be faceted by country.


On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?

```{r}
library(dplyr)
library(tidyr)
library(lubridate)

#a. identify countries have more than 50 reviews.
country50 <- chocolate %>%
  group_by(country_of_bean_origin) %>%
  filter(n() >= 50)
#b. create new variable- 4 groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90%  
country50$cocoa_num <- as.numeric(substr(country50$cocoa_percent, 1, nchar(country50$cocoa_percent)-1))

country50 <- country50 %>% mutate(percgroup = case_when(
        cocoa_num < 60~ 1,
        cocoa_num >= 60 & cocoa_num <70 ~ 2,
        cocoa_num >= 70 & cocoa_num <90 ~ 3,
        cocoa_num >=90 ~ 4
    ))

 ggplot(country50 , aes(x = as.factor (percgroup), y = rating)) +
  geom_boxplot() +
  facet_wrap(~ country_of_bean_origin) +               # Facet the plot by 'country'
  labs(x = "Cocoa Percentage Group",    # Label for x-axis
       y = "Rating",                    # Label for y-axis
       title = "Ratings by Cocoa Percentage Group, Faceted by Country") +
  theme_minimal()   

```





