---
title: "project1"
author: "Peiyu Chen"
date: "2024-09-07"
output: html_document
---
## check package and data


```{r  message=FALSE}

## Install the tidyverse package if you don't have it
if (!require("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
}

## Install the remotes package if you don't have it
if (!require("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
## Install the tidytuesdayR package (from GitHub) if you don't have it
if (!require("tidytuesdayR", quietly = TRUE)) {
    remotes::install_github("dslc-io/tidytuesdayR")
}

##to aviod redownloading data:

library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
  dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
  url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
  chocolate <- readr::read_csv(url_csv)
  
  saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

glimpse(chocolate)

```

## Part 1: Explore data

1. Make a histogram of the rating scores to visualize the overall distribution scores. 
Change the number of bins from default to 10, 15, 20, and 25. 
Pick on the one that you think looks the best.
Explain what the difference is when you changes the number of bins and explain why you picked the one you did. 

```{r message=FALSE}
plot(pressure)
```

2. Consider the countries where the beans originated from. How many reviews come from each country of bean origin?