---
title: "project1"
author: "Peiyu Chen"
date: "2024-09-07"
output: html_document
---


## check package and data


```{r  message=FALSE}

## Install the tidyverse package if you don't have it
if (!require("tidyverse", quietly = TRUE)) {
    install.packages("tidyverse")
}

## Install the remotes package if you don't have it
if (!require("remotes", quietly = TRUE)) {
    install.packages("remotes")
}
## Install the tidytuesdayR package (from GitHub) if you don't have it
if (!require("tidytuesdayR", quietly = TRUE)) {
    remotes::install_github("dslc-io/tidytuesdayR")
}

##to aviod redownloading data:

library("here")
library("tidyverse")

# tests if a directory named "data" exists locally
if (!dir.exists(here("data"))) {
  dir.create(here("data"))
}

# saves data only once (not each time you knit a R Markdown)
if (!file.exists(here("data", "chocolate.RDS"))) {
  url_csv <- "https://raw.githubusercontent.com/rfordatascience/tidytuesday/master/data/2022/2022-01-18/chocolate.csv"
  chocolate <- readr::read_csv(url_csv)
  
  saveRDS(chocolate, file = here("data", "chocolate.RDS"))
}

chocolate <- readRDS(here("data", "chocolate.RDS"))
as_tibble(chocolate)

glimpse(chocolate)

```




## Part 1: Explore data


1. Make a histogram of the rating scores to visualize the overall distribution scores. 
    Change the number of bins from default to 10, 15, 20, and 25. 
    Pick on the one that you think looks the best.Explain what the difference is when you changes the number of bins and explain why you picked the one you did. 

    Solution: In this question, I would like to use 15 bins to demonstrate the distribution of overall rating scores. With 15 bins, the graph looks smoother. When using 20 or 25 bins, there are too many gaps, which makes the graph looks more discrete and makes it harder to observe the distribution pattern.




```{r message=FALSE}
hist(chocolate$rating, main = "#bins=default")
hist(chocolate$rating, breaks=10, main = "#bins=10")
hist(chocolate$rating, breaks=15, main = "#bins=15")
hist(chocolate$rating, breaks=20, main = "#bins =20")
hist(chocolate$rating, breaks=25, main = "#bins =25")

```

   

2. Consider the countries where the beans originated from. How many reviews come from each country of bean origin?

```{r}
beans <- dplyr::filter(chocolate, company_location == country_of_bean_origin)
count(dplyr::group_by(beans,country_of_bean_origin))
```
 
 
 3. What is average rating scores from reviews of chocolate bars that have Ecuador as country_of_bean_origin in this dataset? 
 
 For this same set of reviews, also calculate 
 
 (1) the total number of reviews and 
 
 (2) the standard deviation of the rating scores. 
 
 Your answer should be a new data frame with these three summary statistics in three columns. Label the name of these columns mean, sd, and total.
 

```{r}
ecuador <- chocolate %>% 
  dplyr::group_by(country_of_bean_origin)%>% 
  dplyr::filter(country_of_bean_origin == "Ecuador")

summary <- ecuador %>%
  dplyr::summarise (
    mean = mean(rating),
    sd = sd(rating),
    total = n()
  )
print(summary)
```


4. Which company location makes the best chocolate (or has the highest ratings on average) with beans from Ecuador?


```{r}

maxecuador <- ecuador %>% 
  dplyr::group_by(company_location)%>%
  dplyr::summarise(mean= mean(rating)) %>%
  arrange(desc(mean))%>%
  slice(1)
print(maxecuador)

```

5. Calculate the average rating across all country of origins for beans. Which top 3 countries (for bean origin) have the highest ratings on average?


    Solution: the top three countries are Tobago, China, and Sao Tome & Principe. 
    

```{r}
high3 <- chocolate %>% dplyr::group_by(country_of_bean_origin) %>%
  dplyr::summarise(mean=mean(rating))%>%
  arrange(desc(mean))%>%
  slice(1:3)
print(high3)
```


6. Following up on the previous problem, now remove any countries of bean origins that have less than 10 chocolate bar reviews. Now, which top 3 countries have the highest ratings on average?


    Solution: The top three countries are Solomon Islands, Congo, and Cuba.
    
    

```{r}
high3_10 <- chocolate %>% group_by(country_of_bean_origin) %>%
  dplyr::filter(n() >=10)%>%
  dplyr::summarise(mean=mean(rating))%>%
  arrange(desc(mean))%>%
  slice(1:3)

print(high3_10)
```

       
    
    

7. For this last part, let’s explore the relationship between percent chocolate and ratings.

Use the functions in dplyr, tidyr, and lubridate to perform the following steps to the chocolate dataset:

    a. Identify the countries of bean origin with at least 50 reviews. Remove reviews from countries are not in this list.

    b. Using the variable describing the chocolate percentage for each review, create a new column that groups chocolate percentages into one of four groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90% 
    (Hint: check out the substr() function in base R and the case_when() function from dplyr – see example below).
    
    c. Using the new column described in #2, re-order the factor levels (if needed) to be starting with the smallest percentage group and increasing to the largest percentage group (Hint check out the fct_relevel() function from forcats).
    
    d. For each country, make a set of four side-by-side boxplots plotting the groups on the x-axis and the ratings on the y-axis. These plots should be faceted by country.


On average, which category of chocolate percentage is most highly rated? Do these countries mostly agree or are there disagreements?


      Answer: In the overall boxplot, the second cocoa percentage group, of which the range between 60% to 70%, and the third cocoa percentage group, of which the range between 70% to 90%, have the highest median of rating scores. And most of the countries agree with the overall patterns. 
      


```{r}
library(dplyr)
library(tidyr)
library(lubridate)
library(ggplot2)

#a. identify countries have more than 50 reviews.
country50 <- chocolate %>%
  dplyr::group_by(country_of_bean_origin) %>%
  dplyr::mutate(review_n = n()
  )%>%
  dplyr::filter(review_n >= 50)
  
#b. create new variable- 4 groups: (i) <60%, (ii) >=60 to <70%, (iii) >=70 to <90%, and (iii) >=90%  
country50$cocoa_num <- as.numeric(substr(country50$cocoa_percent, 1, nchar(country50$cocoa_percent)-1))

country50 <- country50 %>% dplyr::mutate(percgroup = case_when(
        cocoa_num < 60 ~ 1,
        cocoa_num >= 60 & cocoa_num <70 ~ 2,
        cocoa_num >= 70 & cocoa_num <90 ~ 3,
        cocoa_num >=90 ~ 4
    ))



ggplot(country50 , aes(x = as.factor (percgroup), y = rating)) +
  geom_boxplot() +
  labs(x = "Cocoa Percentage Group",
       y = "Rating",      
       title = "Ratings by Cocoa Percentage Group, overall") 

 ggplot(country50 , aes(x = as.factor (percgroup), y = rating)) +
  geom_boxplot() +
  facet_wrap(~ country_of_bean_origin) + 
  labs(x = "Cocoa Percentage Group",
       y = "Rating",      
       title = "Ratings by Cocoa Percentage Group, Faceted by Country") 

```

     

## Part 2: Joint dataset together


Tasks


1. Use this dataset it to create a new column called continent in our chocolate dataset1 that contains the continent name for each review where the country of bean origin is.


2. Only keep reviews that have reviews from countries of bean origin with at least 10 reviews.


3. Also, remove the country of bean origin named "Blend".


4. Make a set of violin plots with ratings on the y-axis and continents on the x-axis.


```{r}
library(gapminder)
#view(gapminder)
gapminder1 <- gapminder %>% 
  dplyr::group_by(country)%>% 
  distinct(country, continent) 



newchocol <- chocolate %>% 
  left_join(gapminder1, by = c("country_of_bean_origin" = "country")) %>%
  dplyr::group_by(country_of_bean_origin) %>%
  dplyr::mutate(
    continent = case_when(
      country_of_bean_origin == "Fiji"| country_of_bean_origin == "Papua New Guinea"| country_of_bean_origin== "Vanuatu"|country_of_bean_origin == "Solomon Islands" ~ "Oceania",
      country_of_bean_origin == "Sao Tome"| country_of_bean_origin == "Congo" ~ "Africa",
      is.na(continent) ~ "Americas",
      TRUE ~ continent
    )
  )%>%
  dplyr::filter(n() >= 10 &country_of_bean_origin != "Blend")


unmatch <- newchocol%>%
 dplyr:: filter(is.na(continent))%>%
  dplyr::group_by(country_of_bean_origin)%>%
  distinct(country_of_bean_origin,continent)
unmatch


ggplot(newchocol, aes(x = continent, y = rating)) +
  geom_violin(trim = FALSE) + 
  labs(x = "Continent", 
       y = "Rating",
       title = "Ratings by Continent")



```




## Part 3 Convert wide data into long data

  We are going to create a set of features for us to plot over time. Use the functions in dplyr and tidyr to perform the following steps to the chocolate dataset:

  1. Create a new set of columns titled beans, sugar, cocoa_butter, vanilla, letchin, and salt that contain a 1 or 0 representing whether or not that review for the chocolate bar contained that ingredient (1) or not (0).
  
  
  2. Create a new set of columns titled char_cocoa, char_sweet, char_nutty, char_creamy, char_roasty, char_earthy that contain a 1 or 0 representing whether or not that the most memorable characteristic for the chocolate bar had that word (1) or not (0). 
  
     For example, if the word “sweet” appears in the most_memorable_characteristics, then record a 1, otherwise a 0 for that review in the char_sweet column (Hint: check out str_detect() from the stringr package).
  
  
  3. For each year (i.e. review_date), calculate the mean value in each new column you created across all reviews for that year. (Hint: If all has gone well thus far, you should have a dataset with 16 rows and 13 columns).
  
  
  4. Convert this wide dataset into a long dataset with a new feature and mean_score column.


```{r}
library(stringr)
 chocolate_new <- chocolate %>%
  replace_na(list(ingredients = "")) %>%
  mutate(
    beans = ifelse(str_detect(ingredients, "B"), 1, 0),
    sugar = ifelse(str_detect(ingredients, "\\bS\\b") & !str_detect(ingredients, "S\\*"), 1, 0),     
    cocoa_butter = ifelse(str_detect(ingredients, "C"), 1, 0),
    vanilla = ifelse(str_detect(ingredients, "V"), 1, 0),
    lecithin = ifelse(str_detect(ingredients, "L"), 1, 0), 
    salt = ifelse(str_detect(ingredients, "Sa"), 1, 0)
  )

chocolate_new<- chocolate_new %>%
  mutate(
    char_cocoa = ifelse(str_detect(most_memorable_characteristics, "cocoa"), 1, 0),   # Cocoa
    char_sweet = ifelse(str_detect(most_memorable_characteristics, "sweet"), 1, 0),   # Sweet
    char_nutty = ifelse(str_detect(most_memorable_characteristics, "nutty"), 1, 0),   # Nutty
    char_creamy = ifelse(str_detect(most_memorable_characteristics, "creamy"), 1, 0), # Creamy
    char_roasty = ifelse(str_detect(most_memorable_characteristics, "roasty"), 1, 0), # Roasty
    char_earthy = ifelse(str_detect(most_memorable_characteristics, "earthy"), 1, 0)  # Earthy
  ) 
 
summary_new <- chocolate_new %>%dplyr::group_by(review_date)%>%
  summarise(
    mean_beans = mean(beans, na.rm = TRUE), 
    mean_sugar = mean(sugar, na.rm = TRUE),
    mean_cocoa_butter = mean(cocoa_butter, na.rm = TRUE), 
    mean_vanilla = mean(vanilla, na.rm = TRUE),  
    mean_lecithin = mean(lecithin, na.rm = TRUE), 
    mean_salt = mean(salt, na.rm = TRUE) ,
    mean_char_cocoa = mean(char_cocoa, na.rm = TRUE),
    mean_char_sweet = mean(char_sweet, na.rm = TRUE),
    mean_char_nutty = mean(char_nutty, na.rm = TRUE),
    mean_char_creamy = mean(char_creamy, na.rm = TRUE),
    mean_char_roasty = mean(char_roasty, na.rm = TRUE),
    mean_char_earthy = mean(char_earthy, na.rm = TRUE)
  )

summary_long<- summary_new %>%
  pivot_longer(
    cols = starts_with("mean_"), 
    names_to = "feature_char",
    values_to = "mean_score" 
  )

print(summary_long)
```



## PART 4


```{r}
ggplot(summary_long, aes(x = review_date, y = mean_score)) +
  geom_point(color = "blue") +                             # Observed points
  geom_smooth(method = "loess", se = FALSE, color = "red") + # Smoothed non-linear trend line
    facet_wrap(~ feature_char, scales = "free_y") +               # One plot for each feature
  labs(
    title = "Mean Scores for Chocolate Features",            # Overall title
    x = "Year",                                             
    y = "Mean Score"
  )
```




## PART 5

Tasks


  Using the chocolate dataset (or any of the modified versions you made throughout this assignment or anything else you wish you build upon it):

  
  1. Make the absolute worst plot that you can. You need to customize it in at least 7 ways to make it awful.
 
 
  2. In your document, write 1 - 2 sentences about each different customization you added (using bullets – i.e. there should be at least 7 bullet points each with 1-2 sentences), and how it could be useful for you when you want to make an awesome data visualization.
  
  
    Solution: In this graph, there are several cunstomizations I created:
    
    1. geom_violin and geom_jitter: the worst graph was overplotted. To make the plot clearer and better, I can only keep one plot. 
    
    
    2. the fill color of geom_violin: the color I selected in this worst graph was not beautiful. I can change another set of platter to make the graph more beautiful.
    
    
    3. geom_violin: this is to plot a violin graph and customize the graph in different preference.  
    
    
    4. the label: in this worst graph, I did not make a clear legend label. To edit, I can change the "No legend" to other variabl.
   
   
    5. the background color, and text color: in this worst graph, it was hard to read my text because of the strange set of color I select for background and text. To make the graph better, I can use different pair of color like: plot.background = element_rect(fill = "white", color="white") and axis.text.x = element_text(color = "black")
  
  
    6. The x-axis: in the worst graph, I rotated the x-axis and use the wrong size of text. 
   
   
    7. the legend: in the worst graph, the size of text was too large. T


```{r}
newchocol$cocoa_num <- as.numeric(substr(newchocol$cocoa_percent, 1, nchar(newchocol$cocoa_percent)-1))
newchocol <- newchocol %>% 
  dplyr::mutate(percgroup = case_when(
        cocoa_num < 60 ~ 1,
        cocoa_num >= 60 & cocoa_num <70 ~ 2,
        cocoa_num >= 70 & cocoa_num <90 ~ 3,
        cocoa_num >=90 ~ 4
    ))


newchocol %>% 
  ggplot(aes(
        x = continent,
        y = rating,        
        fill = continent

    )) +
    geom_violin()+
  facet_wrap(~as.factor(percgroup))+
  geom_jitter(width = 0.2, color = "green", height=0.2) +  
  labs(
    x = "X-axis: continents",
    y = "Y-axis: Rating",
    fill = "No Legend"    
  ) +
  theme(
    plot.background = element_rect(fill = "lightyellow"), #light color
    axis.text.x = element_text(angle = 90, size = 36, color = "yellow"), # Hard to read, rotated x-axis text
    axis.text.y = element_text(size = 5), 
    legend.position = "bottom",                             # Clunky legend position
    legend.text = element_text(size = 50)                   # Overly large legend text
  )
    
    
```




## PART 6


```{r}
newchocol %>% 
  ggplot(aes(
        x = continent,
        y = rating,        
        fill = as.factor(percgroup)  # Fill based on percgroup, treat as factor
    )) +
    geom_violin(trim = FALSE, alpha = 0.6, color = "black") +   # Use alpha and black outline
    scale_fill_brewer(palette = "Set3") +                       # Use a color palette for better visuals
    facet_wrap(~as.factor(percgroup))+
    labs(
      title = "Distribution of Ratings by Continent and Percentage Group",  # Clear and informative title
      x = "Continents",                                                    # Improved x-axis label
      y = "Rating",                                                        # Improved y-axis label
      fill = "Percentage of Cocoa Category"                                           # Clarified legend lab
    ) +
    theme_minimal() +                                                      # Clean theme
    theme(
      plot.background = element_rect(fill = "white", color="white", linewidth = 1),
      axis.text.x = element_text(size = 12, color = "black"), 
      axis.text.y = element_text(size = 12), 
      legend.position = "bottom",                             
      legend.text = element_text(size = 12),
      plot.title = element_text(size = 16, face = "bold", hjust = 0.5)  # Centered and bold title
    )
```


    Solution:   In this graph, there are several cunstomizations I edited to make the graph clear:
    
    1. geom_violin and geom_jitter: I only keep "geom_violin"
    
    
    2. the fill color of geom_violin: I change another set of platter by using scale_fill_brewer(palette = "Set3"). 
    
    
    3. geom_violin: I added alpha = 0.6 and color = "black" to make the violin graph slightly transparent and add the black outline. 
    
    
    4. the label: I changed the "No legend" to other variable, like the categories of percentage of cocoa. 
   
   
    5. the background color, and text color: I used a different pair of colors like: plot.background = element_rect(fill = "white", color="white") and axis.text.x = element_text(color = "black") which made the text more clear.
  
  
    6. The x-axis: I deleted the angle=90, and changed from size 36 to size 12.
   
   
    7. the legend: I changed the text size from 36 to 16. 






